---
title: "Decision Tree Modelling using pengiun dataset"
author: "Harris Rizwan"
date: "2023-12-22"
output: html_document
---

Loading data


```{r}

install.packages('caret')
install.packages("rpart.plot")

library(caret)

```



```{r}

data <- read.csv2("E:\\AppDS Semester 1\\IML assignment\\Decision Tree Modelling using pengiun dataset\\PenguinsWithoutMissingValues.csv")

head(data)


```

Removing unnecessary columns for model and performing onehotencoding on Species and Island column


```{r}

df1 = subset(data, select = -c(IndividualID) )


cols_to_encode <- c("Species", "Island")

# Perform one-hot encoding
one_hot_encoded <- model.matrix(~.-1 + ., data = df1[, cols_to_encode])

# Combine the one-hot-encoded matrix with the original dataframe
df1 <- cbind(df1, one_hot_encoded)


# Remove the original categorical columns
df1 <- df1[, !names(df1) %in% cols_to_encode]


df1


```



Making 10 fold k val, setting predictions to true to save our results. 

```{r}
library(caret)

ctrl<- trainControl(method  = "cv", number  = 10, savePredictions = TRUE)

```

Making model and speculating it 

```{r}

model <- train(Gender ~ ., data = df1, method = "rpart", trControl = ctrl)


# Using names() function
variable_names <- names(df1)
print(variable_names)



# Summarize the model
print(model)


```

Visualizing the best model using R part

```{r}

library(rpart.plot)
rpart.plot(model$finalModel)

```


Checking accuracy using confusion Matrix

```{r}
df1$Gender <- factor(df1$Gender)

conf_matrix <- confusionMatrix(predict(model, newdata = df1), df1$Gender)
print(conf_matrix)

```

Predicting Gender

```{r}

df1 <- read.csv2("E:\\AppDS Semester 1\\IML assignment\\Decision Tree Modelling using pengiun dataset\\PenguinsWithoutMissingValues.csv")


cols_to_encode <- c("Species", "Island")

df1 = subset(df1, select = -c(IndividualID) )


# Perform one-hot encoding
one_hot_encoded <- model.matrix(~.-1 + ., data = df1[, cols_to_encode])

# Combine the one-hot-encoded matrix with the original dataframe
df1 <- cbind(df1, one_hot_encoded)


# Remove the original categorical columns
df1 <- df1[, !names(df1) %in% cols_to_encode]

df1

```

Making predictions

```{r}

# Using names() function
variable_names <- names(df1)
print(variable_names)


# Use the trained model to make predictions on the new data
predictions <- predict(model, newdata = df1)

# Print or use the predictions as needed
print(predictions)

```

Adding predicted values to the dataframe that does not have gender variable, we predicted it and now we,re gonna add it to the df. 

```{r}
df1 = subset(data2, select = -c(Gender) )

combined_df <- cbind(df1, predictions)

#renameing predictions to Gender
colnames(combined_df)[colnames(combined_df) == "predictions"] <- "Gender"


head(combined_df)

```

