---
title: "EDA Pengium Dataset"
author: "Harris Rizwan"
date: "2023-12-19"
output: html_document
---

```{r}

install.packages("ggplot2")
install.packages("GGally")
install.packages("dplyr")
install.packages("plotly")
install.packages("psych")
install.packages("doBy")
install.packages("doBy")


library(ggplot2)
library(GGally)
library(dplyr)
library(plotly)
library(psych)
library(doBy)

```

```{r}

# Read CSV file into a data frame
df1 <- read.csv2("E:\\AppDS Semester 1\\IML assignment\\EDA penguins\\PenguinsWithoutMissingValues.csv")
head(df1)

```
We have semicolon (;) as seperator in csv instead of comma(,) . So



Exercise 1. How many observations do we have? How many variables?
Ans:No of colums are 8. 

Code:

```{r}


num_columns <- ncol(df1)
num_columns
```

Exercise 2. What are the names of the observed variables?
Ans: Name of variables are "IndividualID",  "Species",  "Island",  "CulmenLength_mm" "CulmenDepth_mm",  "FlipperLength_mm",  "BodyMass_g",  "Gender"    

Code:

```{r}

variable_names <- names(df1)
variable_names
```

Exercise 3. Provide a summary statistics of the columns?
Ans: 6 variables are right skewed and two variables are left skewed as can be observed under skew variable of detailed_summary

Code:
```{r}


# Use describe() from psych for more detailed summary statistics
detailed_summary <- describe(df1)

# Print the detailed summary statistics
detailed_summary

```

Exercise 4. How many penguins per Species are there?
Ans: No of Pengiuns per Adelie Spcies are 146, No of Pengiuns per Chinstrap  Spcies are 68, No of Pengiuns per Gentoo  Spcies are 119 

Code:

```{r}

library(dplyr)

library(magrittr)


# df1 is my dataframe
penguins_per_species <- df1 %>%
  group_by(Species) %>%
  summarize(Count = n())
penguins_per_species

# i believe this is better way to do it, it is less code :P

species_counts <- table(df1$Species)
species_counts
```


Exercise 5. How many penguins per Species and Gender are there?
Ans: There are total 146 Adelie specie pengiuns from which 73 are female and 73 are male, 
there are total 68 Chinstrap specie pengiuns from which 34 are female and 34 are male, there are total 119 Gentoo specie pengiuns from which 58 are female and 61 are male
Code:

```{r}


# Pengiun per specie and gender
penguins_count <- df1 %>%
  group_by(Species, Gender) %>%
  summarize(Count = n())

penguins_count
```

Exercise 6. Show a BodyMass_g statistics per Species and Gender?
Ans: This observation aligns with a common trend in many bird species where males are often larger than females. The difference in body mass can be attributed to various factors, including reproductive roles, foraging patterns, and environmental conditions

Code:

```{r}

library(dplyr)

# Grouping by Species, and Gender and calculating descriptive statistics
summary_table <- df1 %>%
  group_by(Species, Gender) %>%
  summarize(
    Number_of_Observations = n(),
    Min_BodyMass = min(BodyMass_g	),
    Median_BodyMass = median(BodyMass_g	),
    Mean_BodyMass = mean(BodyMass_g	),
    Max_BodyMass = max(BodyMass_g	)
  )


# View the resulting table
summary_table



```

Exercise 7. Quantitative statistics per Species and Gender?
Ans: For Adelie penguins, both males and females have similar numbers of observations (73 each), but males tend to have a higher mean culmen length, culmen depth, and body mass compared to females

Code:

```{r}


library(dplyr)

# Grouping by Island, Species, and Gender and calculating descriptive statistics
statistics_table <- df1 %>%
  group_by(Species, Gender) %>%
  summarize(
    Number_of_Observations = n(),                
    Mean_CulmenLength_mm = mean(CulmenLength_mm),
    Mean_CulmenDepth_mm = mean(CulmenDepth_mm),
    Mean_FlipperLength_mm = mean(FlipperLength_mm),
    Mean_BodyMass_g = mean(BodyMass_g)  
  )

# View the resulting table
print(statistics_table)


```



Exercise no 8:
Ans:

Code:

```{r}

# Extract numerical variables
numerical_vars <- df1 %>%
  select_if(is.numeric)

# Pairs plot for numerical variables
ggpairs(numerical_vars)

# Pairs plot with color by Species
ggpairs(df1, columns = 4:7, aes(color = Species))

```

Exercise no 9:
Ans:

Code:

```{r}
library(ggplot2)

# Standard Histogram
ggplot(df1, aes(x = CulmenLength_mm)) +
  geom_histogram(binwidth = 1, fill = "skyblue", color = "black") +
  labs(title = "Standard Histogram", x = "Culmen Length (mm)", y = "Frequency")



# Histogram with Different Colors for Different Species
ggplot(df1, aes(x = CulmenLength_mm, fill = Species)) +
  geom_histogram(binwidth = 1, color = "black", position = "identity", alpha = 0.7) +
  labs(title = "Histogram by Species", x = "Culmen Length (mm)", y = "Frequency") +
  scale_fill_manual(values = c("Adelie" = "skyblue", "Chinstrap" = "lightgreen", "Gentoo" = "salmon"))

# Facet Plot with Histograms for Different Species
ggplot(df1, aes(x = CulmenLength_mm, fill = Species)) +
  geom_histogram(binwidth = 1, color = "black", position = "identity", alpha = 0.7) +
  facet_wrap(~Species, scales = "free") +
  labs(title = "Facet Plot by Species", x = "Culmen Length (mm)", y = "Frequency") +
  scale_fill_manual(values = c("Adelie" = "skyblue", "Chinstrap" = "lightgreen", "Gentoo" = "salmon"))

# Facet Plot with Histograms for Different Species and Gender
ggplot(df1, aes(x = CulmenLength_mm, fill = Gender)) +
  geom_histogram(binwidth = 1, color = "black", position = "identity", alpha = 0.7) +
  facet_grid(Species ~ Gender, scales = "free") +
  labs(title = "Facet Plot by Species and Gender", x = "Culmen Length (mm)", y = "Frequency") +
  scale_fill_manual(values = c("MALE" = "skyblue", "FEMALE" = "salmon"))


```
Exercise no 10
Ans:

Code:

```{r}

# Density plot for CulmenLength_mm for different Species, facet_grid by Gender
ggplot(df1, aes(x = CulmenLength_mm, fill = Species)) +
  geom_density(alpha = 0.5) +
  facet_grid(Gender ~ .) +
  labs(title = "Density plot for CulmenLength_mm by Species (facet_grid: Gender)",
       x = "CulmenLength_mm",
       y = "Density")

# Density plot for CulmenLength_mm for different Species, facet_grid by Island
ggplot(df1, aes(x = CulmenLength_mm, fill = Species)) +
  geom_density(alpha = 0.5) +
  facet_grid(Island ~ .) +
  labs(title = "Density plot for CulmenLength_mm by Species (facet_grid: Island)",
       x = "CulmenLength_mm",
       y = "Density")

```

Exercise no 11:

```{r}

# Boxplot for CulmenLength_mm
ggplot(df1, aes(x = "", y = CulmenLength_mm)) +
  geom_boxplot() +
  labs(title = "Boxplot for CulmenLength_mm",
       x = "",
       y = "CulmenLength_mm")

# Boxplot for CulmenLength_mm, color by Species, facet_grid by Gender
ggplot(df1, aes(x = "", y = CulmenLength_mm, fill = Species)) +
  geom_boxplot() +
  facet_grid(Gender ~ .) +
  labs(title = "Boxplot for CulmenLength_mm (facet_grid: Gender, colored by Species)",
       x = "",
       y = "CulmenLength_mm")


```

Exercise no 12:

```{r}

# Scatterplot for CulmenLength_mm and BodyMass_g, color by Species, facet_grid by Gender
ggplot(df1, aes(x = CulmenLength_mm, y = BodyMass_g, color = Species)) +
  geom_point() +
  facet_grid(Gender ~ .) +
  labs(title = "Scatterplot for CulmenLength_mm and BodyMass_g (facet_grid: Gender, colored by Species)",
       x = "CulmenLength_mm",
       y = "BodyMass_g")

```
Exercise no 13:

```{r}

# Create ggplot scatterplot
gg_scatterplot <- ggplot(df1, aes(x = CulmenLength_mm, y = BodyMass_g, color = Species, text = IndividualID)) +
  geom_point() +
  facet_grid(Gender ~ .) +
  labs(title = "Interactive Scatterplot with ggplotly (facet_grid: Gender, colored by Species)",
       x = "CulmenLength_mm",
       y = "BodyMass_g")

# Convert ggplot to ggplotly
ggplotly(gg_scatterplot, tooltip = "text") 


```




